---
title: "rAmChart package"
output:
  html_document:
    toc: true
    theme: flatly
    highlight: tango
---

The package rAmChart allows to draw JavaScript charts wich are by nature interactive.
Its architecture is based on S4 classes and corresponds almost exactly to the
original library [AmCharts][url_amcharts].

We do not focus on the **htmlwidgets** features which make a such API realizable. For more details consult directly <http://www.htmlwidgets.org>.

# Getting started

## pipeR

By loading the package, you will note that it is plainly compatible with other
plot methods since it uses S4 signature for each functionnality.
```{r,echo=FALSE}
library(rAmChart)
```

For the next part of this short introduction, you will always see the symbole
`%>>%` from the package [*pipeR*][home_pipeR]. If you want more details, you can refer to the its tutorial, but in this context we just have to keep in mind that thanks to this the code before is evaluated and passed as the first argument of the function which follow `%>>%` or it can can be any argument by indicating a `.` using `()` or `{}`.

For example:
```{r}
data(iris)
summary(iris, digits = 2)
```
is equivalent to:
```{r}
library(pipeR)
data(iris)
iris %>>% summary( digits = 2 )
```
or even
```{r}
library(pipeR)
data(iris)
iris %>>% ( summary(object = ., digits = 2) )
```

---

## Quick tour:

### Example

Begin with a simple pie chart
```{r, message=FALSE}
library(rAmChart)
amPieChart(theme ="light", titleField = "key", valueField = "value"
) %>>% setDataProvider(data.frame(key = c("FR", "US"), value = c(20,10))
) %>>% setLegend( position = "right"
) %>>% plot
```

### Details...

If you would have seen the same example from a demonstration on <http://www.amcharts.com/demos/>, you'd read a javascript source code like this:
```
{
"type": "pie",
"theme": "light",
"dataProvider": [ {
"country": "FR",
"litres": 20
}, {
"country": "US",
"litres": 10
}],
"valueField": "litres",
"titleField": "country",
}
```

Although the syntax appears very different, it has similarities.
If you observe the result before calling the `plot` method, you see a global list containing each parameter precised in the code:
```{r, message=FALSE}
library(rAmChart)
amPieChart(theme ="light", titleField = "key", valueField = "value"
) %>>% setDataProvider(data.frame(key = c("FR", "US"), value = c(20,10))
)
```

This list can be easily mapped to the previous JavaScript code with the
following tips:

* A JavaScript object corresponds to a list in __R__
* A JavaScript array (a wrapper for object) also corresponds to a list in __R__

This structure and equivalencies are fundamental to construct charts with the package.

The key thing is to imagine the function 
```{r, eval=FALSE}
library(rAmChart)
amPieChart(theme ="light", titleField = "key", valueField = "value")
```
as an initializer of an S4 class  which is a list copntainer for every property.
```{r}
library(rAmChart)
class(amPieChart())
```

* Each element put in the function is a first level property.
* Other complex first level property, like `dataProvider`, has its dedicated function whose the first argument is always an object of S4 class **AmChart** (or **AmStock**).

### Customization...

Currently you can draw any chart from:
* <http://docs.amcharts.com/3/javascriptcharts>
* <http://docs.amcharts.com/3/javascriptstockchart>

That means you can customize you chart in various ways !

For instance, we can move the default linked label *JS chart by amCharts*, which must be kept for using the library without license.
We also can switch the theme:

```{r, message=FALSE}
library(rAmChart)
amPieChart(theme ="chalk", titleField = "key", valueField = "value",
           creditsPosition = "bottom-left"
) %>>% setDataProvider(data.frame(key = c("FR", "US"), value = c(20,10))
) %>>% setLegend( position = "right"
) %>>% plot
```

# Few demos

## Basics

### amXYChart

```{r, message=FALSE}
library(rAmChart)
amXYChart(theme = "default", startDuration = 0.5, marginLeft = 46, marginBottom = 35
) %>>% setDataProvider(data.frame( y = c(10,5,-10,-6,15,13,1),  x = c(14,3,8,5,-4,1,6),
                                   value = c(59,50,19,65,92,8,16), y2 = c(-5,-15,-4,-5,-10,-2,0),
                                   x2 = c(-3,-8,6,-6,-8,0,-3), value2 = c(44,12,35,168,102,41,16))
) %>>% addValueAxes(position = "bottom", axisAlpha = 0
) %>>% addValueAxes(minMaxMultiplier = 1.2, position = "left", axisAlpha = 0
) %>>% addGraph(balloonText = "x:<b>[[x]]</b> y:<b>[[y]]</b><br>value:<b>[[value]]</b>",
                bullet = "circle", bulletBorderAlpha = 0.2, bulletAlpha = 0.8, lineAlpha=0,
                fillAlphas = 0, valueField = "value", xField = "x",yField = "y", maxBulletSize = 100
) %>>% addGraph(balloonText = "x:<b>[[x]]</b>y:<b>[[y]]</b><br>value:<b>[[value]]</b>",
                bullet="diamond",bulletBorderAlpha=0.2, bulletAlpha = 0.8, lineAlpha = 0,
                fillAlphas = 0,valueField = "value2", xField = "x2", yField = "y2",
                maxBulletSize = 100
) %>>% setExport %>>% plot
```

### gaugeChart

```{r, message=FALSE}
library(rAmChart)
amAngularGaugeChart(theme = "light"
) %>>% addArrow(value = 100
) %>>% addAxe(gaugeAxis (bottomText = "100 km/h", endValue = 220, valueInterval = 10
) %>>% addBand(color = "#00CC00", endValue = 90, startValue = 0
) %>>% addBand(color = "#ffac29", endValue = 130, startValue = 90
) %>>% addBand(color = "#ea3838", endValue = 220, startValue = 130, innerRadius = "95%")
) %>>% setExport %>>% plot
```

### funnelChart

```{r, message=FALSE}
library(rAmChart)
amFunnelChart(theme = "light", neckHeight = "30%", neckWidth = "40%", titleField = "title",
              valueField = "value", creditsPosition = "bottom-left", 
              dataProvider = data.frame(title = c("Website visits", "Downloads"),
                                        value = c(300, 123))
) %>>% plot
```

## Elaborated

### Column and line mix

See <http://www.amcharts.com/demos/column-and-line-mix/>

```{r, message=FALSE}
library(rAmChart)
amSerialChart(addClassNames = TRUE, theme = "light", autoMargins = FALSE, marginLeft = 30,
              marginRight = 8, marginTop = 10, marginBottom = 26, startDuration = 1,
              categoryField = "year"
) %>>% setDataProvider(data.frame(year  =  c(2010, 2011, 2012, 2013, 2014),
                                  income  =  c(20.4, 20.6, 24.3, 21.5, 22.3), 
                                  expenses  =  c(31.5, 32.4, 25.6, 22.6, 24.9),
                                  dashLengthColumn = c(NA, NA, NA, NA, 5), 
                                  alpha = c(NA, NA, NA, NA, 0.2), 
                                  additional = c("", "", "", "", "(projection)"))
) %>>% setBalloon(adjustBorderColor = FALSE, horizontalPadding = 10,
                  verticalPadding = 8,color = "#ffffff"
) %>>% addValueAxes(axisAlpha = 0, position = "left"
) %>>% addGraph(alphaField = "alpha", fillAlphas = 1, title = "Income",
                type = "column", valueField = "income",
                balloonText = "<span style = 'font-size:12px;'>[[title]] in [[category]]:<br><span style = 'font-size:20px;'>[[value]]</span> [[additional]]</span>", 
                 dashLengthField = "dashLengthColumn"
) %>>% addGraph(id = "graph2", bullet = "round", lineThickness = 3, bulletSize = 7,
                balloonText = "<span style = 'font-size:12px;'>[[title]] in [[category]]:<br><span style = 'font-size:20px;'>[[value]]</span> [[additional]]</span>", 
                bulletBorderAlpha = 1, bulletColor = "#FFFFFF",
                useLineColorForBulletBorder = TRUE, bulletBorderThickness = 3, 
                fillAlphas = 0, lineAlpha = 1, title = "Expenses", valueField = "expenses"
) %>>% setCategoryAxis(gridPosition = "start", axisAlpha = 0, tickLength = 0
) %>>% setExport %>>% plot
```

### Column with rotated series

See <http://www.amcharts.com/demos/exporting-chart-to-image/>

```{r, message=FALSE}
library(rAmChart)
amSerialChart(theme = "dark", startDuration = 2, categoryField = "country", depth3D = 40,
               angle = 30
) %>>% setDataProvider(data.frame(
  country = c("USA", "China", "Japan", "Germany", "UK", "France", "India", "Spain", "Netherlands", "Russia"), 
  visits = c(3025, 1882, 1809, 1322, 1122, 1114, 984, 711, 665, 580), 
  color = c("#FF0F00", "#FF6600", "#FF9E01", "#FCD202", "#F8FF01", "#B0DE09", "#04D215", "#0D8ECF", "#0D52D1", "#2A0CD0"))
) %>>% addGraph(balloonText = "<b>[[category]]: [[value]]</b>", fillColorsField = "color", 
                 fillAlphas = 0.85, lineAlpha = 0.1, type = "column", valueField = "visits"
) %>>% setChartCursor(categoryBalloonEnabled = FALSE, cursorAlpha = 0, zoomable = FALSE
) %>>% setCategoryAxis(gridPosition = "start", labelRotation = 45, axisAlpha = 0, gridAlpha = 0
) %>>% setExport %>>% plot
```

### Stacked areas

See <http://www.amcharts.com/demos/stacked-area/>

```{r, message=FALSE}
library(rAmChart)
amSerialChart(theme = "light", marginRight = 30, plotAreaBorderAlpha = 0,
               categoryField = "year"
) %>>% setDataProvider(data.frame(
  year  =  1994:2012, 
  cars  =  c(15, 13, 14, 15, 16, 18, 19, 22, 24, 20, 24, 25, 26, 35, 36, 37, 38, 39, 40), 
  motorcycles  =  c(15, 13, 14, 15, 16, 18, 19, 22, 24, 20, 24, 25, 26, 35, 36, 37, 38, 39, 40), 
  bicycles  =  c(15, 13, 14, 15, 16, 18, 19, 22, 24, 20, 24, 25, 26, 35, 36, 37, 38, 39, 40))
) %>>% setLegend(equalWidths = FALSE, periodValueText = "total: [[value.sum]]", position = "top", valueAlign = "left", valueWidth = 100
) %>>% addValueAxes(stackType = "regular", gridAlpha = 0.07, position = "left",
                    title = "Traffic incidents"
) %>>% addGraph(balloonText = "<img src = 'http://www.amcharts.com/lib/3/images/car.png' style = 'vertical-align:bottom; margin-right: 10px; width:28px; height:21px;'><span style = 'font-size:14px; color:#000000;'><b>[[value]]</b></span>", 
                 fillAlphas = 0.6, hidden = TRUE, lineAlpha = 0.4, title = "Cars",
                valueField = "cars"
) %>>% addGraph(balloonText = "<img src = 'http://www.amcharts.com/lib/3/images/motorcycle.png' style = 'vertical-align:bottom; margin-right: 10px; width:28px; height:21px;'><span style = 'font-size:14px; color:#000000;'><b>[[value]]</b></span>", 
                 fillAlphas = 0.6, lineAlpha = 0.4, title = "Motorcycles", valueField = "motorcycles"
) %>>% addGraph(balloonText = "<img src = 'http://www.amcharts.com/lib/3/images/bicycle.png' style = 'vertical-align:bottom; margin-right: 10px; width:28px; height:21px;'><span style = 'font-size:14px; color:#000000;'><b>[[value]]</b></span>", 
                 fillAlphas = 0.6, lineAlpha = 0.4, title = "Bicycles", valueField = "bicycles"
) %>>% setChartCursor(cursorAlpha = 0
) %>>% setCategoryAxis(startOnAxis = TRUE, axisColor = "#DADADA", gridAlpha = 0.07
) %>>% setGuides(list(guide(category = "2001", toCategory = "2003", lineColor = "#CC0000",
                            lineAlpha = 1, fillAlpha = 0.2, fillColor = "#CC0000", dashLength = 2,
                            inside = TRUE, labelRotation = 90,
                            label = "fines for speeding increased"),
                      guide(category = "2007", lineColor = "#CC0000", lineAlpha = 1,
                            dashLength = 2, inside = TRUE, labelRotation = 90,
                            label =  "motorcycle fee introduced"))
) %>>% setExport %>>% plot
```

### Waterfall chart

See <http://www.amcharts.com/demos/waterfall-chart/>

```{r, message=FALSE}
library(rAmChart)
amSerialChart(theme = "light", startDuration = 1, columnWidth = 0.6, categoryField = "name"
) %>>% setDataProvider( data.frame(
  name = c("Income A", "Income B", "Total Income", "Expenses A", "Expenses B", "Revenue"), 
  open  =  c(0, 11.13, 0, 12.92, 8.64, 0), 
  close =  c(11.13, 15.81, 15.81, 15.81, 12.92, 8.64), 
  color  =  c("#54cb6a", "#54cb6a", "#169b2f", "#cc4b48", "#cc4b48", "#1c8ceb"), 
  balloonValue  =  c(11.13, 4.68, 15.81, 2.89, 4.24, 11.13) )
) %>>% addValueAxes(axisAlpha = 0, gridAlpha = 0.1, position = "left"
) %>>% addGraph(balloonText = "<span style = 'color:[[color]]'>[[category]]</span><br><b>$[[balloonValue]] Mln</b>", 
                 colorField = "color", fillAlphas = 0.8, labelText = "$[[balloonValue]]",
                lineColor = "#BBBBBB", openField = "open", type = "column", valueField = "close"
) %>>% addTrendLine(dashLength = 3, finalCategory = "Income B",
                    finalValue = 11.13, initialCategory = "Income A",
                    initialValue = 11.13, lineColor = "#888888"
) %>>% addTrendLine(dashLength = 3, finalCategory = "Expenses A", finalValue = 15.81, initialCategory = "Income B", initialValue = 15.81, lineColor = "#888888"
) %>>% addTrendLine(dashLength = 3, finalCategory = "Expenses A",
                    finalValue = 15.81, initialCategory = "Income B",
                    initialValue = 15.81, lineColor = "#888888"
) %>>% addTrendLine(dashLength = 3, finalCategory = "Expenses B",
                    finalValue = 12.92, initialCategory = "Expenses A",
                    initialValue = 12.92, lineColor = "#888888"
) %>>% addTrendLine(dashLength = 3, finalCategory = "Revenue", finalValue = 8.64,
                    initialCategory = "Expenses B", initialValue = 8.64, lineColor = "#888888"
) %>>% setCategoryAxis(gridPosition = "start", axisAlpha = 0, gridAlpha = 0.1
) %>>% setExport %>>% plot
```

### Candle stick chart

See <http://www.amcharts.com/demos/candlestick-chart/>

```{r, message=FALSE}
library(rAmChart)
amSerialChart(categoryField = "date", theme = "light", dataDateFormat = "YYYY-MM-DD"
) %>>% setDataProvider(data.frame (date = paste0(2011, "-", 8:11, "-", replicate(4, 1:30)),
                                     open = round(runif(length(paste0(2011, "-", 8:11, "-", replicate(4, 1:30))), 130, 140), 2), 
                                     high = round(runif(length(paste0(2011, "-", 8:11, "-", replicate(4, 1:30))), 140, 150), 2),
                                     low = round(runif(length(paste0(2011, "-", 8:11, "-", replicate(4, 1:30))), 120, 130), 2),
                                     close = round(runif(length(paste0(2011, "-", 8:11, "-", replicate(4, 1:30))), 120, 150), 2))
) %>>% addValueAxes(position = "left"
) %>>% addGraph(id = "g1",
                 balloonText = "Open : <b>[[open]]</b><br>Low =<b>[[low]]</b><br>High =<b>[[high]]</b><br>Close =<b>[[close]]</b><br>",
                 closeField = "close", fillColors = "#7f8da9", highField = "high",
                 lineColor = "#7f8da9", lineAlpha = 1, lowField = "low",
                 fillAlphas = "0.9",  negativeFillColors = "#db4c3c", negativeLineColor = "#db4c3c",
                 openField = "open", title = "Price:", type = "candlestick", valueField = "close"
) %>>% setChartCursor(valueLineEnabled = TRUE, valueLineBalloonEnabled = TRUE
) %>>% addListener("rendered", "function(event) { event.chart.zoomToIndexes(event.chart.dataProvider.length - 10, event.chart.dataProvider.length - 1) }"
) %>>% setExport(position = "bottom-right") %>>% plot
```

### Stock chart with multiple data sets

See <http://www.amcharts.com/demos/multiple-data-sets/>

```{r, message=FALSE}
library(rAmChart)
firstDate <- Sys.Date()
chartData1 <- as.data.frame(t(sapply(0:20, FUN = function(i)
{
  date <- format(firstDate + i, "%m/%d/%Y")
  a <- round(runif(1) * (40 + i)) + 100 + i
  b <- round(runif(1) * (1000 + i)) + 500 + i * 2
  c(date = date, value = a,  volume = b)
})))

chartData2 <- as.data.frame(t(sapply(0:20, FUN = function(i)
{
  date <- format(firstDate + i, "%m/%d/%Y")
  a <- round(runif(1) * (100 + i)) + 200 + i
  b <- round(runif(1) * (1000 + i)) + 600 + i * 2
  c(date = date, value = a,  volume = b)
})))

chartData3 <- as.data.frame(t(sapply(0:20, FUN = function(i)
{
  date <- format(firstDate + i, "%m/%d/%Y")
  a <- round(runif(1) * (100 + i)) + 200 + i
  b <- round(runif(1) * (1000 + i)) + 600 + i * 2
  c(date = date, value = a,  volume = b)
})))

chartData4 <- as.data.frame(t(sapply(0:20, FUN = function(i)
{
  date <- format(firstDate + i, "%m/%d/%Y")
  a <- round(runif(1) * (100 + i)) + 200 + i
  b <- round(runif(1) * (1000 + i)) + 600 + i * 2
  c(date = date, value = a,  volume = b)
})))

amStockChart(theme = "light"
) %>>% addDataSet(dataSet(title = "first data set", categoryField = "date",
                            dataProvider = chartData1) %>>%
                     addFieldMapping(fromField = "value", toField = "value") %>>%
                     addFieldMapping(fromField = "volume", toField = "volume")
) %>>% addDataSet(dataSet(title = "second data set", categoryField = "date",
                            dataProvider = chartData2) %>>%
                     addFieldMapping(fromField = "value", toField = "value") %>>%
                     addFieldMapping(fromField = "volume", toField = "volume")
) %>>% addDataSet(dataSet(title = "third data set", categoryField = "date",
                            dataProvider = chartData3) %>>%
                     addFieldMapping(fromField = "value", toField = "value") %>>%
                     addFieldMapping(fromField = "volume", toField = "volume")
) %>>% addDataSet(dataSet(title = "fourth data set", categoryField = "date",
                            dataProvider = chartData4) %>>%
                     addFieldMapping(fromField = "value", toField = "value") %>>%
                     addFieldMapping(fromField = "volume", toField = "volume")
) %>>% addPanel(stockPanel(showCategoryAxis = FALSE, title = "Value", percentHeight = 70) %>>%
                   addStockGraph(id = "g1", valueField = "value", comparable = TRUE,
                                 compareField = "value", balloonText = "[[title]] =<b>[[value]]</b>",
                                 compareGraphBalloonText = "[[title]] =<b>[[value]]</b>"
                  ) %>>% setStockLegend(periodValueTextComparing = "[[percents.value.close]]%",
                                          periodValueTextRegular = "[[value.close]]")
) %>>% addPanel(stockPanel(title = "Volume", percentHeight = 30) %>>%
                   addStockGraph(valueField = "volume", type = "column",
                                 fillAlphas = 1) %>>%
                   setStockLegend(periodValueTextRegular = "[[value.close]]")
) %>>% setChartScrollbarSettings(graph = "g1"
) %>>% setChartCursorSettings(valueBalloonsEnabled = TRUE, fullWidth = TRUE,
                               cursorAlpha = 0.1, valueLineBalloonEnabled = TRUE,
                               valueLineEnabled = TRUE, valueLineAlpha = 0.5
) %>>% setPeriodSelector(periodSelector(position = "left") %>>%
                            addPeriod(period = "MM", selected = TRUE, count = 1, label = "1 month") %>>%
                            addPeriod(period = "MAX", label = "MAX")
) %>>% setDataSetSelector(position = "left"
) %>>% setPanelsSettings(recalculateToPercents = FALSE
) %>>% plot
```

### Stock chart with dates

See <http://www.amcharts.com/demos/multiple-data-sets/>

```{r, message=FALSE}
library(rAmChart)
start <- as.POSIXct("01-01-2015", format = "%d-%m-%Y")
end <- as.POSIXct("31-12-2015", format = "%d-%m-%Y")
period <- seq.POSIXt(from = start, to = end, by = "10 min")
n <- length(period)
periodTemp <- seq.POSIXt(from = start, to = end, by = "3 hour")
nTemp <- length(periodTemp)
### Generate mesures ----
charge <- rnorm(n, mean = 500, sd= 200)
charge[ which(charge < 0) ] <- rnorm(length(which(charge < 0)), mean = 200, sd = 10)
temp <- rnorm(nTemp, mean = 15, sd = 10)
dtCharge <- data.table::data.table(charge, date = period)
setkey(dtCharge, date)
dtTemp <- data.table::data.table(temperature = temp, date = periodTemp)
setkey(dtTemp , date)
dp <- dtTemp[dtCharge]
dp[ , date := format(date, "%m-%d-%Y %H:%M:%S")]

nTest <- 10000
sample <- 1:nTest
amStockChart(theme = "default" , dataDateFormat = "MM-DD-YYYY JJ:NN:SS"
) %>>% addDataSet(dataSet(title = "Courbe de charge", categoryField = "date") %>>%
                    setDataProvider(dp[ sample, ], keepNA = FALSE) %>>%
                    addFieldMapping(fromField = "charge", toField = "charge") %>>%
                    addFieldMapping(fromField = "temperature", toField = "temperature") %>>%
                    addStockEvent( date = "01-02-2015 23:00:00", type = "sign", graph ="g1",
                                    text = "I am a stockEvent", description = "I am a property of a DataSet")
) %>>% addPanel(stockPanel(showCategoryAxis = FALSE, title = "Charge", percentHeight = 70) %>>%
                   addStockGraph(id = "g1", valueField = "charge", comparable = TRUE,
                                 compareField = "charge", balloonText = "[[title]] =<b>[[value]]</b>",
                                 compareGraphBalloonText = "[[title]] =<b>[[value]]</b>"
                  ) %>>% setStockLegend(periodValueTextComparing = "[[percents.value.close]]%",
                                          periodValueTextRegular = "[[value.close]]")
) %>>% addPanel(stockPanel(title = "Temperature", percentHeight = 30) %>>%
                   addStockGraph(valueField = "temperature", fillAlphas = 1) %>>%
                   setStockLegend(periodValueTextRegular = "[[value.close]]")
) %>>% setChartScrollbarSettings(graph = "g1"
) %>>% setChartCursorSettings(valueBalloonsEnabled = TRUE, fullWidth = TRUE,
                               cursorAlpha = 0.1, valueLineBalloonEnabled = TRUE,
                               valueLineEnabled = TRUE, valueLineAlpha = 0.5
) %>>% setPeriodSelector(periodSelector(position = "left") %>>%
                            addPeriod(period = "MM", selected = TRUE, count = 1, label = "1 month") %>>%
                            addPeriod(period = "MAX", label = "MAX")
) %>>% setDataSetSelector(position = "left"
) %>>% setCategoryAxesSettings(minPeriod = "ss"
) %>>% setPanelsSettings(recalculateToPercents = FALSE
) %>>% plot

```

[url_amcharts]: www.amcharts.com
[home_pipeR]:http://renkun.me/pipeR-tutorial/